### 2023-02-08 article3.qmd
update

all_floats <- rbind(float_6904240, float_6904241, float_4903634, float_1902578)

all_floats <- all_floats |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))

argo_map <- all_floats |> select(WMO = wmo, cycle, lat, lon) |> drop_na(lat) |> dplyr::group_by(WMO, cycle) |> mutate(WMO = factor(WMO)) |>
  dplyr::summarise(lat = mean(lat), lon = mean(lon))

@@

world_data <- map_data("world")
w <- list(geom_polygon(aes(long, lat, group=group), data=world_data, fill="grey70"), coord_quickmap())

@@

# remove bad data from float 6904241
float_6904241_slope <- float_6904241 |> filter(cycle < 20)
# create dataframe for all drifting data (for all floats)
all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)

@@

cflux$date <- as_date(cflux$min_time)

# # add levels
# cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))

@@

# add level for order
#uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
# uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))

@@

compute_flux <- function(A, b, PSD){ # PSD = Particle Size Distribution
  # for classes sizes between 0.102 mm and 1.5 mm
  #mid_ESD <- c(0.1164512, 0.1463600, 0.1843915, 0.2325200, 0.2933257, 0.3691650, 0.4650400, 0.5860455, 0.7385533, 0.9280422, 1.1705684, 1.4795321) # With Rainer's code (paruvpy)
  # mid_ESD <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
  #              1.16170742, 1.46365963) # With Lionel's script
  mid_ESD <- c(0.23051195, 0.29042685,
               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
               1.16170742, 1.46365963) # With Lionel's script
    # mid_ESD <- c(0.29042685,
    #            0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
    #            1.16170742, 1.46365963) # With Lionel's script
  exp <- A * mid_ESD ** b
  estimated_flux <- rowSums(t(t(PSD)*exp))
  return(estimated_flux)
}

parking_data <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
parking_data <- parking_data |> filter(PhaseName == 'PAR') |> drop_na(NP_Size_102) |> select(park_depth, all_of(lpm_class), dates = juld, park_depth, wmo, cycle) |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
parking_spectra <- parking_data |> dplyr::select(NP_Size_203:NP_Size_1290)

parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo) #|> summarize(median_guidi_flux = median(guidi_flux))

@@

# https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles
# ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = small_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
#   geom_errorbar(aes(x = date, y = guidi_flux),
#                       stat = 'summary',
#                       fun.min = function(z) {quantile(z, 0.25)},
#                       fun.max = function(z) {quantile(z, 0.75)},
#                       fun = median, colour = 'red', width = 5) +
#   geom_line(aes(x = date, y = guidi_flux),
#                       stat = 'summary',
#                       fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') +
#   geom_line(data = comparison_flux_ost_uvp, aes(x = date, y = big_flux), linetype = 'dotdash') + theme_bw()

ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
  geom_errorbar(aes(x = date, y = guidi_flux),
                      stat = 'summary',
                      fun.min = function(z) {quantile(z, 0.25)},
                      fun.max = function(z) {quantile(z, 0.75)},
                      fun = median, colour = 'red', width = 5) +
  geom_line(aes(x = date, y = guidi_flux),
                      stat = 'summary',
                      fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') + theme_bw()

correlations <- comparison_flux_ost_uvp |> select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> group_by(wmo, park_depth, cycle) |> mutate(median_guidi_flux = median(guidi_flux)) |> group_by(park_depth, wmo) |> summarise(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))

correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_big, cor_total))
gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Big flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
  cols_label(wmo = 'WMO', `cor_small_200 m` = '200 m',
             `cor_small_500 m` = '500 m',
             `cor_small_1000 m` = '1000 m',
             `cor_big_200 m` = '200 m',
             `cor_big_500 m` = '500 m',
             `cor_big_1000 m` = '1000 m',
             `cor_total_200 m` = '200 m',
             `cor_total_500 m` = '500 m',
             `cor_total_1000 m` = '1000 m')

@@

-   faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -\> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
-   discuter des corrélations
-   science: on voit le système latent, on a raté le bloom
-   rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire) -> permet d'utiliser le graphique de l'UVP en vertical -> hop une justification en plus
-   rajouter dans les méthodes les produits satellites qu'on utilise
-   différentes dynamiques post-bloom
-   phénologie du bloom printanier

### Which particles drive the flux

```{r, flux model 2, include = FALSE}

flux_model <- function(boosting_rounds, which_flux){

    # select which flux to build the model
    if(which_flux == 'small'){
      cflux2 <- cflux |> mutate(flux = small_flux) 
    }else if(which_flux == 'big'){
      cflux2 <- cflux |> mutate(flux = big_flux)
    }else{
      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
    }

@@

  # training data
  tmp <- data_model |> select(flux, all_of(lpm_class))

@@

  return(list(features_importance = features_importance, model = model, best_tree = best))

# small_model <- flux_model(3000, 'small')
# big_model <- flux_model(3000, 'big')
# total_model <- flux_model(3000, 'total')
# 
# save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data')

#load(file = 'DATA/MODEL/model_xgboost_data')

@@

#### On profiling mode

Cette partie devrait venir dans la discussion si on part du principe que la première partie est liée à la méthodologie

@fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes. In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 650 m and 1050 m. Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August. It can be seen among all size classes that the water column is being drained proportionally to the class size.

```{r, uvp_data on the vertical}
#| label: fig-uvp-vertical-profiles
#| fig-cap: Normalized concentrations of particles for all UVP6 vertical profiles. Plots were arranged to respect the deployment of each couple of floats (e.g. the first two rows of plots are related to co-deployed floats 6904240 and 4903634). The colorbar depicts the log$_{10}$ of normalized concentrations. Grey areas indicate that particles were not measured in the size range of interest.

# particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))
# 
# particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
# #particles_bis$size <- factor(particles_bis$size, levels = lpm_class)
# 
# # rel_conc
# particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))

# order floats by couple of deployment
# particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))

# stats on the local maximum
# test <- particles_bis |> filter(depth > 500) |> group_by(wmo, cycle, size) |> # find depth of maximum relative abundance below 500 m between June and September
#   summarize(index_max = which.max(rel_conc), max_depth = depth[index_max], date = unique(juld)) |> 
#   filter(size < 512, date < '2022-10-01', date > '2022-05-31') |> # compute mean depth for each class size for each float
#   group_by(wmo, size) |> summarize(mean_depth = mean(max_depth))

# plot concentration
# particles_bis |> filter(depth < 200) |>
#   ggplot() + facet_grid(wmo~size) + theme_bw() +
#   geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
#   #scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
#   scale_colour_viridis(option = 'H') +
#   scale_y_reverse() +
#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
#   labs(x = 'Month', y = 'Depth (m)') +
#   geom_path(data = particles_bis, aes(juld, MLD), colour = "white")

# ## plot absolute concentrations
# up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
#   #group_by(size) |>
#   ggplot() + facet_grid(wmo~size) + theme_bw() +
#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
#   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
#   #scale_colour_viridis(option = 'A', limits = c(-2,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
#   #scale_colour_viridis(option = 'H') +
#   scale_y_reverse() +
#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
#   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)') #+
#   #geom_hline(yintercept = c(600, 1100), linetype = 'dotted')
#   #theme(legend.title = element_text(angle = -90))
#   #guides(fill = guide_legend(title.theme = element_text(size = 200)))
#   #+ theme(legend.position = 'none')
#   #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
#   #theme(legend.position = 'none') #+
#   #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
#   #guides(colour = guide_legend(title.position = "bottom"))
# 
# #Normalized concentrations (#/L)
# 
# up_part
# 
# down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
#   #group_by(size) |>
#   ggplot() + facet_grid(wmo~size) + theme_bw() +
#   geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
#   #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
#   scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
#   #scale_colour_viridis(option = 'H') +
#   scale_y_reverse() +
#   scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
#   #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
#   labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
# 
# down_part
```

@@

 + aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies

@@

@@

@@

tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
  # compute concentration anomaly at 500 and 1000 m
  mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
  select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
  mutate(size = as.character(size)) |>
  mutate(rev_div_anomaly = 1/div_anomaly) |> 
  #filter(is.finite(rev_div_anomaly)) |>
  mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
  ungroup()

@@

# plot_heatmap <- function(float){
#   y <- unique(tmp$size)[1:12]
#   tt <- tmp |> filter(wmo == float, size %in% y) 
#   x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
#   data <- expand.grid(X=x, Y=y)
#   #data$Z <- tt$anom
#   data$Z <- tt$diff_anomaly
#   data$Z <- tt$rev_div_anomaly
# 
#   p <- ggplot(data, aes(X, Y, fill= Z)) + 
#     geom_tile() + scale_fill_viridis() +
#     geom_text(data = data, aes(X, Y, label = Z)) +
#     labs(title = float)
#   
#   return(p)
# }
# 
# p1 <- plot_heatmap(4903634)
# p2 <- plot_heatmap(6904240)
# p3 <- plot_heatmap(1902578)
# p4 <- plot_heatmap(6904241)
# 
# (p1 | p2) / 
# (p3 | p4)

>>>>

# combine all data
all_floats <- rbind(float_6904240, float_6904241, float_4903634, float_1902578)

# add levels for some field
all_floats <- all_floats |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))

# summarize localization of floats
argo_map <- all_floats |> select(WMO = wmo, cycle, lat, lon) |> drop_na(lat) |> dplyr::group_by(WMO, cycle) |> mutate(WMO = factor(WMO)) |>
  dplyr::summarise(lat = mean(lat), lon = mean(lon))

@@

# map positions of floats
world_data <- map_data("world")
w <- list(geom_polygon(aes(long, lat, group=group), data=world_data, fill="grey70"), coord_quickmap())

@@

# remove bad data from float 6904241
float_6904241_slope <- float_6904241 |> filter(cycle < 20)
# create dataframe for all drifting data (for all floats)
all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle) |>
   mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
          park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))

@@

cflux$date <- as_date(cflux$min_time)
```

@@

@@

compute_flux <- function(A, b, PSD){ # PSD = Particle Size Distribution
  # for classes sizes between 0.102 mm and 1.5 mm
  #mid_ESD <- c(0.1164512, 0.1463600, 0.1843915, 0.2325200, 0.2933257, 0.3691650, 0.4650400, 0.5860455, 0.7385533, 0.9280422, 1.1705684, 1.4795321) # With Rainer's code (paruvpy)
  # mid_ESD <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
  #              1.16170742, 1.46365963) # With Lionel's script
  # mid_ESD <- c(0.23051195, 0.29042685,
  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
  #              1.16170742, 1.46365963) # With Lionel's script
  #mid_ESD <- c(0.287555212,0.362129811,0.455929819,0.57466512,0.724144323,0.91063714,1.147083258,1.450068964)
  mid_ESD <- c(0.29042685,
               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
               1.16170742, 1.46365963) # With Lionel's script
  exp <- A * mid_ESD ** b
  estimated_flux <- rowSums(t(t(PSD)*exp))
  return(estimated_flux)
}

parking_data <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
parking_data <- parking_data |> filter(PhaseName == 'PAR') |> drop_na(NP_Size_102) |> select(park_depth, all_of(lpm_class), dates = juld, park_depth, wmo, cycle) |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
parking_spectra <- parking_data |> dplyr::select(NP_Size_256:NP_Size_1290)

parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo)

@@

# https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles
ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
  geom_errorbar(aes(x = date, y = guidi_flux),
                      stat = 'summary',
                      fun.min = function(z) {quantile(z, 0.25)},
                      fun.max = function(z) {quantile(z, 0.75)},
                      fun = median, colour = 'red', width = 5) +
  geom_line(aes(x = date, y = guidi_flux),
                      stat = 'summary',
                      fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') + theme_bw()

@@

```{r, correlations between OST and UVP carbon fluxes}
#| label: tbl-correlation
#| tbl-cap: TODO
correlations <- comparison_flux_ost_uvp |> select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> group_by(wmo, park_depth, cycle) |> mutate(median_guidi_flux = median(guidi_flux)) |> group_by(park_depth, wmo) |> summarise(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))

correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_big, cor_total))

gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Big flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
  cols_label(wmo = 'WMO', `cor_small_200 m` = '200 m',
             `cor_small_500 m` = '500 m',
             `cor_small_1000 m` = '1000 m',
             `cor_big_200 m` = '200 m',
             `cor_big_500 m` = '500 m',
             `cor_big_1000 m` = '1000 m',
             `cor_total_200 m` = '200 m',
             `cor_total_500 m` = '500 m',
             `cor_total_1000 m` = '1000 m')

Figure @fig-comparison-ost-uvp shows a comparison of carbon fluxes estimated with the OST ($F_{total} = F_{small} + F_{big}$) and the UVP6 (median of all UVP measurements while the float was drifting) at all drifting depths. Overall, carbon fluxes estimated with the coefficients ($A$, $b$) of @Guidi2008-vp, are lower than the ones measured by the OST. Without additional measurements (e.g. co-deployed sediment traps), it is however impossible to decipher which measurements is closest to the truth. Nonetheless, the trends measured by both instruments are very well correlated at all drifting depths for floats 6904240 and 6904241 with correlations comprised between `r min(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])` and `r max(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,8:10])`. @tbl-correlation also shows that the correlation between UVP-derived carbon fluxes and $F_{small}$ and $F_{big}$ is `r mean(as.vector(as.matrix(correlations[correlations$wmo %in% c(6904240, 4903634),][1:2,2:7])))` on average. We calculated lower correlations with the other two floats, especially the float 1902578 at 200 m, where correlations between the UVP-derived carbon fluxes and all fluxes (small, big and total) are close to 0.

Floats equipped with both a transmissometer and a UVP6 represent an opportunity to look at the influence of parts of the size spectrum on $F_{small}$ and $F_{big}$. For this, we decided to build a regression model using XGBoost, a gradient boosting method [REF], using biovolumes measured by the UVP to predict the slope of $F_{small}$ and $F_{big}$ measured by the OST.

# compute biovolumes for mid size bin (ESD)
# DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)

# DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
#               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
#               1.16170742, 1.46365963, 1.84409558, 2.32341484)
# 
# # DSE obtained with Rainer's code
# DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
#                 0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
#                 0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
#                 1.4795320531800742,1.8636618038457486,2.3449457391257185)

DSE <- c(0.29042685,0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779, 1.16170742, 1.46365963)

# compute biovolume
get_biovolume <- function(DSE){
  r <- DSE/2
  biovolume <- (4*pi/3)*r**3
  return(biovolume)
}

# get biovolume for each DSE (equivalent to say for each size bin)
biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3

# spectra at parking depth
uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))

# convert abundance to biovolume
spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')

# add data to initial abundance spectra
uvp_spectra <- cbind(uvp_spectra, spectra_biovolume)

# recompute slope based on Estapa's formula
cflux <- cflux |> mutate(small_slope = (small_flux/633)**(1/0.77), big_slope = (big_flux/633)**(1/0.77))

## build model
flux_model <- function(boosting_rounds, which_slope){

  # select which flux to build the model
  if(which_slope == 'small'){
    cflux2 <- cflux |> mutate(flux = log(small_slope)) |> filter(is.finite(flux))
  }else{
    cflux2 <- cflux |> mutate(flux = log(big_slope)) |> filter(is.finite(flux)) # because log(0) creates non finite value -> bug
  }

@@

  # training data
  tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))

@@

  return(list(features_importance = features_importance, model = model, best_tree = best))
}

# small_model <- flux_model(3000, 'small')
# big_model <- flux_model(3000, 'big')
# save(small_model, big_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data_slope')

load(file = 'DATA/MODEL/model_xgboost_data_slope')

# # get spectra
# uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |> 
#   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
# 
# flux_model <- function(boosting_rounds, which_flux){
#     
#     # select which flux to build the model
#     if(which_flux == 'small'){
#       cflux2 <- cflux |> mutate(flux = small_flux) 
#     }else if(which_flux == 'big'){
#       cflux2 <- cflux |> mutate(flux = big_flux)
#     }else{
#       cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
#     }
#     
#     # merge data for training
#     data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
#     
#   # training data
#   tmp <- data_model |> select(flux, all_of(lpm_class))
#   
#   set.seed(101)
#   # model
#   model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
#   cv.folds=5,
#   n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
#   bag.fraction=0.5,
#   n.cores=1
#   )
#    
#   # get best model and thus best round number
#   best <- gbm.perf(model)
#   #print(best)
#   
#   # get features importance
#   features_importance <- summary(model, n.trees=best, plotit = FALSE)
#   
#   
#   return(list(features_importance = features_importance, model = model, best_tree = best))
#    
# }
#   
# # small_model <- flux_model(3000, 'small')
# # big_model <- flux_model(3000, 'big')
# # total_model <- flux_model(3000, 'total')
# # 
# # save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data') 
#   
# load(file = 'DATA/MODEL/model_xgboost_data')

@@

model_predict <- function(which_slope, mod, best_tree_number){

    # convert abundance to biovolume
    spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes)))
    colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')

    # add data to initial abundance spectra
    uvp_spectra <- cbind(uvp_spectra, spectra_biovolume)

    # select which flux to build the model
    if(which_slope == 'small'){
      cflux2 <- cflux |> mutate(flux = small_slope)
    }else{
      cflux2 <- cflux |> mutate(flux = big_slope)
    }

    # # training data
    # tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))

    # predictions on mean spectra values
    uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = log(unique(flux)),
                                                                                   NP_Biovolume_102 = mean(NP_Biovolume_102),
                                                                                   NP_Biovolume_128 = mean(NP_Biovolume_128),
                                                                                   NP_Biovolume_161 = mean(NP_Biovolume_161),
                                                                                   NP_Biovolume_203 = mean(NP_Biovolume_203),
                                                                                   NP_Biovolume_256 = mean(NP_Biovolume_256),
                                                                                   NP_Biovolume_323 = mean(NP_Biovolume_323),
                                                                                   NP_Biovolume_406 = mean(NP_Biovolume_406),
                                                                                   NP_Biovolume_512 = mean(NP_Biovolume_512),
                                                                                   NP_Biovolume_645 = mean(NP_Biovolume_645),
                                                                                   NP_Biovolume_813 = mean(NP_Biovolume_813),
                                                                                   NP_Biovolume_1020 = mean(NP_Biovolume_1020),
                                                                                   NP_Biovolume_1290 = mean(NP_Biovolume_1290),
                                                                                   NP_Biovolume_1630 = mean(NP_Biovolume_1630),
                                                                                   NP_Biovolume_2050 = mean(NP_Biovolume_2050)) |> ungroup()

# predictions
# pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
# pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux) 
# all_predictions <- cbind(pred_small, pred_big)
# vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions_slopes.csv')
all_predictions <- vroom('DATA/MODEL/all_predictions_slopes.csv')

# model_predict <- function(which_flux, mod, best_tree_number){
# 
#     # get spectra
#     uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |>
#   drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))
# 
#     # select which flux to build the model
#     if(which_flux == 'small'){
#       cflux2 <- cflux |> mutate(flux = small_flux)
#     }else if(which_flux == 'big'){
#       cflux2 <- cflux |> mutate(flux = big_flux)
#     }else{
#       cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
#     }
# 
#     # merge data for training
#     data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)
# 
#   # predictions on mean spectra values
#   uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(flux),
#                                                                                    NP_Size_102 = mean(NP_Size_102),
#                                                                                    NP_Size_128 = mean(NP_Size_128),
#                                                                                    NP_Size_161 = mean(NP_Size_161),
#                                                                                    NP_Size_203 = mean(NP_Size_203),
#                                                                                    NP_Size_256 = mean(NP_Size_256),
#                                                                                    NP_Size_323 = mean(NP_Size_323),
#                                                                                    NP_Size_406 = mean(NP_Size_406),
#                                                                                    NP_Size_512 = mean(NP_Size_512),
#                                                                                    NP_Size_645 = mean(NP_Size_645),
#                                                                                    NP_Size_813 = mean(NP_Size_813),
#                                                                                    NP_Size_1020 = mean(NP_Size_1020),
#                                                                                    NP_Size_1290 = mean(NP_Size_1290),
#                                                                                    NP_Size_1630 = mean(NP_Size_1630),
#                                                                                    NP_Size_2050 = mean(NP_Size_2050)) |> ungroup()
# 
# 
#   # prediction at 1000 m
#   #spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))
# 
#   # prediction
#   pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)
# 
#   uvp_spectra_average$pred_flux <- pred_flux
# 
#   return(uvp_spectra_average)
# }
# 
# # pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
# # pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux)
# # pred_total <- model_predict('total', total_model$model, total_model$best_tree) |> select(true_total = true_flux, pred_total = pred_flux)
# #
# # all_predictions <- cbind(pred_small, pred_big, pred_total) |> mutate(pred_total2 = pred_small + pred_big, anom = pred_total - pred_total2)
# #
# # vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions.csv')
# all_predictions <- vroom('DATA/MODEL/all_predictions.csv')

```{r, plot model predictions}

p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + theme_bw() 
round(R2_Score(y_true = all_predictions$true_small, y_pred = all_predictions$pred_small), 2)

p_big <- ggplot(all_predictions |> filter(is.finite(true_big))) + geom_point(aes(true_big, pred_big)) + theme_bw()
all_predictions |> filter(is.finite(true_big)) |> summarize(round(R2_Score(true_big, pred_big)), 2)

p_small / p_big

-   science: on voit le système latent, on a raté le bloom
-   rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire) -\> permet d'utiliser le graphique de l'UVP en vertical -\> hop une justification en plus
-   rajouter dans les méthodes les produits satellites qu'on utilise
-   différentes dynamiques post-bloom
-   phénologie du bloom printanier
-

@@

-   aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies

@@

```
```{r, sinking speed values}

gt(all_sinking_speed)

@@

J'ai enlevé le truc du spectre inversé à 500 et 1000 m car difficile de conclure -> mais comme on est dans la meme masse d'eau, on pourrait le garder, why not

@@

# tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
#   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
#   # compute concentration anomaly at 500 and 1000 m
#   mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
#   select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
#   mutate(size = as.character(size)) |>
#   mutate(rev_div_anomaly = 1/div_anomaly) |> 
#   #filter(is.finite(rev_div_anomaly)) |>
#   mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
#   ungroup()

@@
